---
title: "OpenCovid_Slack"
output: html_notebook
---


Parsing json objects retrieved using Slack API. Objective is to process the data and create dataframes of converstion threads, replies and emoticon responses for network analysis.


```{r}

library("jsonlite")
library("igraph")
library("visNetwork")
library("stringr")



path = "OpenCOVID19 Initiative Slack export Mar 3 2020 - Jan 22 2021" # Path to the root folder with the API   extracted Slack data
folders = list.files(path)
  
users_json = read_json(path = paste(path,"/users.json", sep = ""))
channels_json = read_json(path = paste(path,"/channels.json", sep = ""))

```

Create Users Dataframe

```{r}


users_df = data.frame()

for (i in 1:length(users_json))
{
  email = ""
  
  if (length(users_json[[i]]$profile$email)>0)
  {
    email = users_json[[i]]$profile$email
  }
  
    if (!users_json[[i]]$deleted)
  {  
    invit = users_json[[i]]$is_invited_user
    invit[is.null(invit)] = FALSE
    
    users_df = rbind(users_df, data.frame(id = users_json[[i]]$id, real_name = users_json[[i]]$real_name, timezone  = users_json[[i]]$tz, is_deleted = FALSE, is_admin = users_json[[i]]$is_admin, is_owner = users_json[[i]]$is_owner, is_primary_owner = users_json[[i]]$is_primary_owner, is_restricted = users_json[[i]]$is_restricted, is_ultra_restricted = users_json[[i]]$is_ultra_restricted, is_bot = users_json[[i]]$is_bot, is_app_user = users_json[[i]]$is_app_user, is_invited_user = invit, updated = users_json[[i]]$updated, title = users_json[[i]]$profile$title, image_url = users_json[[i]]$profile$image_512, email = email))
    }
  
  else
  {
    users_df = rbind(users_df, data.frame(id = users_json[[i]]$id, real_name = users_json[[i]]$profile$real_name, timezone = "", is_deleted = TRUE, is_admin = "", is_owner = "", is_primary_owner = "", is_restricted = "",  is_ultra_restricted = "", is_bot = users_json[[i]]$is_bot, is_app_user = users_json[[i]]$is_app_user, is_invited_user = "", updated = users_json[[i]]$updated, title = users_json[[i]]$profile$title, image_url = users_json[[i]]$profile$image_512,  email = email))
  }
  
}


write.csv(users_df, "data_files/users.csv")

```


Create Channels Dataframe

```{r}

channels_df = data.frame()
channels_aux = data.frame()

for (i in 1:length(channels_json))
{
  channels_df = rbind(channels_df, data.frame(id = channels_json[[i]]$id, name = channels_json[[i]]$name, created = channels_json[[i]]$created, creator = channels_json[[i]]$creator, is_archived = channels_json[[i]]$is_archived, is_general = channels_json[[i]]$is_general, members = toString(unlist(channels_json[[i]]$members))))
  
  users = c()
  
  for (j in channels_json[[i]]$pins)
  {
    users = c(users, j$user)
  }
  
  users = toString(users)
  
  channels_aux = rbind(channels_aux, data.frame(id = channels_json[[i]]$id, topic = channels_json[[i]]$topic$value, topic_creator = channels_json[[i]]$topic$creator, topic_ts = channels_json[[i]]$topic$last_set, purpose = channels_json[[i]]$purpose$value, purpose_creator = channels_json[[i]]$purpose$creator, purpose_ts = channels_json[[i]]$purpose$last_set, pinned = users))
  
}

write.csv(channels_df, "data_files/channels.csv")
write.csv(channels_aux, "data_files/channels_desc.csv")

```

Function to process Slack JSON


Construct Network


1. Channel - Identify users not in users.json

```{r}

process_Slack_JSON = function(file_path)
{
  content = read_json(path = file_path)
  list = c()
  users_list = data.frame()
  
  for (i in 1:length(content))
  {
    
      if (!is.null(content[[i]]$subtype)) 
      {
        var = content[[i]]$subtype
        #print(var)
        
        if (var == "bot_message")
        {
          list = c(list,i)
        }
      }
    
    if (!(i %in% list))
    {
      if (!(content[[i]]$user %in% users_df$id))
      {
              users_list = rbind(users_list, data.frame(id = content[[i]]$user))
      }
    }
  
  }
  
  #print(list)
  return(users_list)
}

```


Process texts for mentions

```{r}

process_texts_mentions = function(file_path, folder)
{
  content = read_json(path = file_path)
  df = data.frame()
  df_list = data.frame()
  lis = list()
  
  for (i in 1:length(content))
  {
    
    list = c()
    
    if (is.null(content[[i]]$subtype))
    {
      list = c(list, str_extract_all(content[[i]]$text, regex("<@[:alnum:]+>"))[[1]])
      list = c(list, str_extract_all(content[[i]]$text, regex("<![:alnum:]+>"))[[1]])
    }
    
    #print(str_extract_all(content[[i]]$text, regex("<![:alnum:]+>")))
    cl_list = list()
    
    for (j in list)
    {
        val = str_replace(j,"<","")
        val = str_replace(val,">","")
        val = str_replace(val,"@","")
        val = str_replace(val,"!","")
        cl_list = append(cl_list, val)
        df = rbind(df, data.frame(from = content[[i]]$user, to = val, timestamp = content[[i]]$ts, channel = folder)) 
    }
    #df_list = rbind(df_list, data.frame(from = content[[i]]$user, to = toString(cl_list), timestamp = content[[i]]$ts, channel = folder))
  }
  
 
  lis[[1]] = df
  #lis[[2]] = df_list
  
  return(lis)
}

```

***************

Process text for emoticon responses - texts - and channel joins

```{r}
process_emoticon_responses = function(file_path,folder)
{
  content = read_json(path = file_path)
  df = data.frame()
  df1 = data.frame()
  df2 = data.frame()
  df3 = data.frame()
  df_list = list()
  
  for (i in 1:length(content))
  {
    if (!is.null(content[[i]]$reactions))
    {
      for (j in 1:length(content[[i]]$reactions))
      {
        if (is.null(content[[i]]$subtype))
          {#print(content[[i]])
          df = rbind(df, data.frame(reaction_by = unlist(content[[i]]$reactions[[j]]$users), source_message_from = content[[i]]$user, source_message_timestamp = content[[i]]$ts, reaction_name = content[[i]]$reactions[[j]]$name,  channel = folder))}
      }
    }
    
    if (is.null(content[[i]]$subtype))
    {
        df1 = rbind(df1, data.frame(source_message_from = content[[i]]$user, source_message_timestamp = content[[i]]$ts, message_size = nchar(content[[i]]$text), channel = folder))
        
        
        if (!is.null(content[[i]]$thread_ts) & is.null(content[[i]]$parent_user_id) & !is.null(content[[i]]$reply_count))
        {
          df3 = rbind(df3, data.frame(user = content[[i]]$user, thread_timestamp = content[[i]]$thread_ts, reply_count = content[[i]]$reply_count, reply_user_count =  content[[i]]$reply_users_count, channel = folder))
        }
    }
    
    if (!is.null(content[[i]]$subtype))
    {
      if(content[[i]]$subtype == "channel_join")
        df2 = rbind(df2, data.frame(user = content[[i]]$user, source_message_timestamp = content[[i]]$ts, channel = folder))
    }
    
      
  }
  
  df_list[[1]] = df
  df_list[[2]] = df1
  df_list[[3]] = df2
  df_list[[4]] = df3
  
  return(df_list)
}
```

Workspace Joins

```{r}

workspace_joins = function(fol_path)
{
  files = list.files(fol_path)
  df = data.frame()
  
  for (j in files)
  {
    content = read_json(path = paste(fol_path,"/",j,sep=""))
    #print(paste(fol_path,"/",j,sep=""))
    #print(j)
    
    for (i in 1:length(content))
    {
      if (!is.null(content[[i]]$subtype))
      {
        if(content[[i]]$subtype == "channel_join")
        {  
          df = rbind(df, data.frame(user = content[[i]]$user, timestamp = content[[i]]$ts, mode = "cj"))
        }
        
        else if(content[[i]]$subtype == "bot_message")
        {
          list = str_extract_all(content[[i]]$text, regex("<@[:alnum:]+>"))[[1]]  
          #print(list)
          
          
          for (k in list)
          {
            val = str_replace(k,"<","")
            val = str_replace(val,">","")
            val = str_replace(val,"@","")
            val = str_replace(val,"!","")
            df = rbind(df, data.frame(user = val, timestamp = content[[i]]$ts, mode = "gb")) 
          }
          
        }
          
      }
    }
  }
  
  return(df)
  
}

```



```{r}

df_extra_users = data.frame()
df_text_mentions = data.frame()
df_emoticon_responses = data.frame()
df_texts = data.frame()
df_channel_join = data.frame()
df_text_ment_list = data.frame()
df_threads = data.frame()

df_ws_joins = workspace_joins(paste(path,"/","prgm-introductions", sep = ""))

for (i in folders[-grep(".json",folders)])
{
    files = list.files(paste(path,"/",i, sep = ""))   
    #temp = c()
    
    for (j in files)
    {
        #df_extra_users = rbind(df_extra_users, process_Slack_JSON(paste(path,"/",i,"/",j,sep="")))
        temp1 = process_texts_mentions(file_path = paste(path,"/",i,"/",j,sep=""), folder = i)
        df_text_mentions = rbind(df_text_mentions, temp1[[1]])
        #df_text_ment_list = rbind(df_text_ment_list, temp1[[2]])
        temp =  process_emoticon_responses(file_path = paste(path,"/",i,"/",j,sep=""), folder = i)
        df_emoticon_responses = rbind(df_emoticon_responses, temp[[1]])
        df_texts = rbind(df_texts, temp[[2]])
        df_channel_join = rbind(df_channel_join, temp[[3]])
        df_threads = rbind(df_threads, temp[[4]])
    }
    
}

df_extra_users = unique(df_extra_users)

```

```{r}
write.csv(df_ws_joins, "data_files/ws_joins.csv")
#write.csv(df_extra_users, "data_files/extra_users.csv")
write.csv(df_emoticon_responses, "data_files/emoticon_responses.csv")
write.csv(df_text_mentions, "data_files/text_mentions.csv")
write.csv(df_channel_join, "data_files/channel_join.csv")
write.csv(df_texts, "data_files/texts_channel.csv")
write.csv(df_threads, "data_files/threads_channel.csv")
```

*******************************************************************************************


