---
title: "Slack Data to HIN Format"
output: html_notebook
---

The objective to convert the data export from a Slack workspace to a Heterogenous Information Network (HIN) format.

The data export is a series of JSON files. Here functions are presented to process them and create the HIN.

Format Description: [To be modified with new insights]

1. Entities (Nodes):
  a. Users
  b. Channels
  c. Posts/Messages
  
2. Relations (Edges):
  a. Users - Channel: "has_joined", "has_created", "has_pinned", "created_topic", "created_purpose" [Done]
  b. Users - Post: "replied_to", "reacted_to", "has_created" [Done]
  c. Users - Users: "has_mentioned" [Done]
  d. Posts - Channel: "belongs_to" [Done]
  e. Posts - User: "is_mentioning" [Done]
  
  
Create a Master Dataframe. This dataframe is of format for each edge (interaction)

Source_ID Source_Type Destination_ID Destination_Type Timestamp Relation  
  
```{r setup}

#require("knitr")
#opts_knit$set(root.dir = '~/metastudy')


```


```{r}

library("jsonlite")
library("stringr")


path = "OpenCOVID19 Initiative Slack export Mar 2 2020 - Jul 27 2020" # Path to the root folder with the API   extracted Slack data
folders = list.files(path)

users_json = read_json(path = paste(path,"/users.json", sep = ""))  #Read users metadata file
channels_json = read_json(path = paste(path,"/channels.json", sep = "")) #Read Channels metadata file

interactions = data.frame() #The Final HIN dataframe



```

Process Users Data - [Haven't stored email ID/Skype or other info. Would require when mapping b/w Slack and JOGL] **

```{r}

users_df = data.frame()

for (i in 1:length(users_json))
{
  email = ""
  
  if (length(users_json[[i]]$profile$email)>0)
  {
    email = users_json[[i]]$profile$email
  }
  
    if (!users_json[[i]]$deleted)
  {  
    invit = users_json[[i]]$is_invited_user
    invit[is.null(invit)] = F
    
    users_df = rbind(users_df, data.frame(id = users_json[[i]]$id, real_name = users_json[[i]]$real_name, timezone  = users_json[[i]]$tz, is_deleted = F, is_admin = users_json[[i]]$is_admin, is_owner = users_json[[i]]$is_owner, is_primary_owner = users_json[[i]]$is_primary_owner, is_restricted = users_json[[i]]$is_restricted, is_ultra_restricted = users_json[[i]]$is_ultra_restricted, is_bot = users_json[[i]]$is_bot, is_app_user = users_json[[i]]$is_app_user, is_invited_user = invit, updated = users_json[[i]]$updated, title = users_json[[i]]$profile$title, image_url = users_json[[i]]$profile$image_512, email = email))
    }
  
  else
  {
    users_df = rbind(users_df, data.frame(id = users_json[[i]]$id, real_name = users_json[[i]]$profile$real_name, timezone = "", is_deleted = T, is_admin = "", is_owner = "", is_primary_owner = "", is_restricted = "",  is_ultra_restricted = "", is_bot = users_json[[i]]$is_bot, is_app_user = users_json[[i]]$is_app_user, is_invited_user = "", updated = users_json[[i]]$updated, title = users_json[[i]]$profile$title, image_url = users_json[[i]]$profile$image_512,  email = email))
  }
  
}

```

Process Channel Data & Add user-channel interactions like "has_pinned", "created_topic", "created_purpose" and "created_channel"

```{r}
channels_df = data.frame()
pins = data.frame()

for (i in 1:length(channels_json))
{
  channels_df = rbind(channels_df, data.frame(id = channels_json[[i]]$id, name = channels_json[[i]]$name, creation = channels_json[[i]]$created, topic = channels_json[[i]]$topic$value, purpose = channels_json[[i]]$purpose$value, is_archived = channels_json[[i]]$is_archived, is_general = channels_json[[i]]$is_general))
  
  interactions = rbind(interactions, data.frame(source_id = channels_json[[i]]$creator, source_type = "user", destination_id = channels_json[[i]]$id, destination_type = "channel", timestamp =channels_json[[i]]$created, relation = "has_created"))
  
  users = c()
  
  for (j in channels_json[[i]]$pins)
  {
    users = c(users, j$user)
    
    #interactions = rbind(interactions, data.frame(source_id = j$user, source_type = "user", destination_id = channels_json[[i]]$id, destination_type = "channel", timestamp = j$created, relation = "created_pin"))
    
    pins = rbind(pins, data.frame(source_id = j$user, id = j$id, timestamp = j$created, channel = channels_json[[i]]$id, type = j$type))
  }
  
  users = toString(users)
  
   if (!channels_json[[i]]$topic$value == "")
  {
    interactions = rbind(interactions, data.frame(source_id = channels_json[[i]]$topic$creator, source_type = "user", destination_id = channels_json[[i]]$id, destination_type = "channel", timestamp =channels_json[[i]]$topic$last_set, relation = "created_topic"))
   }
  
  if (!channels_json[[i]]$purpose$value == "")
  {
     interactions = rbind(interactions, data.frame(source_id = channels_json[[i]]$purpose$creator, source_type = "user", destination_id = channels_json[[i]]$id, destination_type = "channel", timestamp =  channels_json[[i]]$purpose$last_set, relation = "created_purpose"))
  }
  
}

```


Process Texts Data

First, some functions to process JSON data

Text mentions: User - User interaction ("has_mentioned") and Post - User ("is_mentioning")

```{r}

process_JSON = function(file_path, folder)
{
  content = read_json(path = file_path)
  interaction = data.frame()
  posts = data.frame()
  
  name = strsplit(strsplit(file_path,"/")[[1]][3], ".json")[[1]]
  
      
    #Deleted/Edited Texts
    
    edited = F
    deleted = F
    
    if (length(content[[i]]$edited)>0)
      edited = T
    
    if (!is.null(content[[i]]$subtype))
    {
      if (content[[i]]$subtype == "tombstone")
        deleted = T
    }
  
  for (i in 1:length(content))
  {
    
      #Process attachments and Files
      
      files = 0
      attachments = 0
      a_list = c()
      f_list = c()
      
      if (length(content[[i]]$attachments)>0)
      {
        attachments = length(content[[i]]$attachments)
        for (j in 1:attachments)
        {
          a_list = c(a_list, content[[i]]$attachments[[j]]$original_url)  
        }
      }
      
      if (length(content[[i]]$files)>0)
      {
        files = length(content[[i]]$files)
        for (j in 1:files)
        {
          f_list = c(f_list, content[[i]]$files[[j]]$id)  
        }
      }
      
      #print(a_list)
      #print(f_list)
      
      posts = rbind(posts, data.frame(id = paste(folder,"_", name,"_", i, sep = ""), text = nchar(content[[i]]$text), timestamp = content[[i]]$ts, is_edited = edited , is_deleted = deleted, files = files, attachments = attachments, a_list = toString(a_list)))

      
    
    list = c()
    
    if (is.null(content[[i]]$subtype) || content[[i]]$subtype == "tombstone")
    {
      
      #Get Users who are mentioned in text
      
      list = c(list, str_extract_all(content[[i]]$text, regex("<@[:alnum:]+>"))[[1]])
      list = c(list, str_extract_all(content[[i]]$text, regex("<![:alnum:]+>"))[[1]])
      
      if(!is.null(content[[i]]$reactions))
      {
        
        #Reacted to a post
        
        for(j in 1:length(content[[i]]$reactions))
        {
           interaction = rbind(interaction, data.frame(source_id = unlist(content[[i]]$reactions[[j]]$users), source_type = "user", destination_id = paste(folder,"_", name,"_", i, sep = ""), destination_type = "message", timestamp = content[[i]]$ts, relation = "reacted_to"))
        }
        
      }
      
      #Created a post
      
      interaction = rbind(interaction, data.frame(source_id = content[[i]]$user, source_type = "user", destination_id = paste(folder,"_", name,"_", i, sep = ""), destination_type = "message", timestamp = content[[i]]$ts, relation = "has_created"))
      
      #Post belongs to 
      
      interaction = rbind(interaction, data.frame(source_id = paste(folder,"_", name,"_", i, sep = ""), source_type = "message", destination_id = channels_df$id[channels_df$name == folder], destination_type = "channel", timestamp = content[[i]]$ts, relation = "belongs_to"))
      
      #User pinned a message
        
        if (content[[i]]$ts %in% pins$id) #If Pin is a message
        {
          interaction = rbind(interaction, data.frame(source_id = pins$source_id[pins$id == content[[i]]$ts], source_type = "user", destination_id = paste(folder,"_", name,"_", i, sep = ""), destination_type = "message", timestamp = as.character(pins$timestamp[pins$id == content[[i]]$ts]), relation = "has_pinned"))
        }
      
        fls = pins$id[pins$id %in% f_list]
        
        if (length(fls)>0) #If Pin is a File
        {
          
          for (k in fls)
          {
            interaction = rbind(interaction, data.frame(source_id = pins$source_id[pins$id ==  k], source_type = "user", destination_id = paste(folder,"_", name,"_", i, sep = ""), destination_type = "message", timestamp = as.character(pins$timestamp[pins$id == k]), relation = "has_pinned"))
          }
          
        }
      
      #Replied messages
      
      if (!is.null(content[[i]]$thread_ts) & is.null(content[[i]]$parent_user_id) & !is.null(content[[i]]$reply_count))
        {
        
        #User replied to post - in a chain of replies - the parent message is considered as the message replied to
        
          for (j in 1:length(content[[i]]$replies))
          {
            interaction = rbind(interaction, data.frame(source_id = content[[i]]$replies[[j]]$user, source_type = "user", destination_id = paste(folder,"_", name,"_", i, sep = ""), destination_type = "message",  timestamp = content[[i]]$replies[[j]]$ts, relation = "replied_to"))
          }
      }
      
    }
    

    
    #print(content[[i]]$subtype)

  
    
    for (j in list)
    {
        val = str_replace(j,"<","")
        val = str_replace(val,">","")
        val = str_replace(val,"@","")
        val = str_replace(val,"!","")
        interaction = rbind(interaction, data.frame(source_id = content[[i]]$user, source_type = "user", destination_id = val, destination_type = "user", timestamp = content[[i]]$ts, relation = "has_mentioned"))
        
        interaction = rbind(interaction, data.frame(source_id = paste(folder,"_", name,"_", i, sep = ""), source_type = "message", destination_id = val, destination_type = "user", timestamp = content[[i]]$ts, relation = "is_mentioning"))
        
    }
    
  
      if (!is.null(content[[i]]$subtype))
      {
        if(content[[i]]$subtype == "channel_join")
        {  
          #print(content[[i]])
          interaction = rbind(interaction, data.frame(source_id = content[[i]]$user, source_type = "user", destination_id = channels_df$id[channels_df$name == folder], destination_type = "channel", timestamp = content[[i]]$ts, relation = "has_joined"))
          
          if (length(content[[i]]$inviter)>0)
          {
            interaction = rbind(interaction, data.frame(source_id = content[[i]]$inviter, source_type = "user", destination_id = content[[i]]$user, destination_type = "user", timestamp = content[[i]]$ts, relation = "has_invited"))
            
            interaction = rbind(interaction, data.frame(source_id = content[[i]]$inviter, source_type = "user", destination_id = channels_df$id[channels_df$name == folder], destination_type = "channel", timestamp = content[[i]]$ts, relation = "invited_to"))
            
          }
          
          
        }
        
        else if(content[[i]]$subtype == "bot_message")
        {
          list = str_extract_all(content[[i]]$text, regex("<@[:alnum:]+>"))[[1]]  #Processed for the "GreetBot" in the Workspace
          
          for (k in list)
          {
            val = str_replace(k,"<","")
            val = str_replace(val,">","")
            val = str_replace(val,"@","")
            val = str_replace(val,"!","")
            #print(val)
            interaction = rbind(interaction, data.frame(source_id = val, source_type = "user", destination_id = channels_df$id[channels_df$name == folder], destination_type = "channel", timestamp = content[[i]]$ts, relation = "has_joined"))
          }
          
        }
          
      }   
    
    
  }
  
  return(list(interaction,posts))
}


```

Call these functions

```{r}
posts = data.frame()


for (i in folders[-grep(".json",folders)])
{
    files = list.files(paste(path,"/",i, sep = ""))   
    
    for (j in files)
    {
        temp = process_JSON(file_path = paste(path,"/",i,"/",j,sep=""), folder = i)
        interactions = rbind(interactions, temp[[1]])
        posts = rbind(posts, temp[[2]])
    }
    
}

```

```{r}

write.csv(interactions, "HIN_edges.csv")
write.csv(users_df, "users.csv")
write.csv(posts, "posts.csv")

```

Convert to JSON (edges)

```{r}
colnames(interactions) = c("StartObject","StartGroup","EndObject","EndGroup","timestamp","Relation")

#process as a big composite list

temp = data.frame(timestamp = interactions$timestamp)
new_int = interactions[,c("Relation","StartGroup", "StartObject", "EndGroup", "EndObject")]
new_int$Attributes = temp

#l = list(Relation = as.list(interactions$Relation), StartGroup = as.list(interactions$StartGroup), StartObject = as.list(interactions$StartObject), EndGroup = as.list(interactions$EndGroup), EndObject = as.list(interactions$EndObject), Attributes = list(timestamp = as.list(interactions$timestamp)))

#intjson = toJSON(new_int, dataframe = "rows")

write_json(new_int, paste("HIN_edges_","Jul_27_2020",".json"), pretty = T)
```

Nodes

```{r}

write_json(users_df, paste("Users","Jul_27_2020",".json"), pretty = T)
write_json(channels_df, paste("Channels","Jul_27_2020",".json"), pretty = T)
write_json(posts, paste("Posts","Jul_27_2020",".json"), pretty = T)

```


*********************************************************************


